---
title: "Take Home Exam"
author: "Michael Zhang"
output: word_document
---
#Book Problems

##Chapter 2, #10
a)
```{r}
library(MASS)
```
There are 506 rows and 14 columns. The rows represent observations (housing values in Boston suburbs), and the columns contain features (e.g. median house value, per capita crime rate, etc.).

b)
```{r}
pairs(Boston)
```
We see a fairly strong negative correlation between lstat and medv. We also see a fairly strong positive correlation between rm and medv and a similarly negative correlation between rm and lstat. Additionally, we see a negative correlation between dis and lstat. There are other correlations between the predictors, but these are some of the strongest that we see from the pairwise scatterplots.

c)
```{r}
plot(Boston$age, Boston$crim)
plot(Boston$dis, Boston$crim)
plot(Boston$rad, Boston$crim)
plot(Boston$tax, Boston$crim)
plot(Boston$ptratio, Boston$crim)
```
Yes, there are predictors associated with crime rate. 
* The older the home, the more crime
* The farther away from Boston employment centers, the less crime
* The more accessible to radial highways, the more crime
* The higher the property tax, the more crime
* The higher the student-teacher ratio, the more crime

d)
```{r}
hist(Boston$crim, 25)
median(Boston$crim)
range(Boston$crim)
quantile(Boston$crim, 0.95)

hist(Boston$tax, 25)
median(Boston$tax)
range(Boston$tax)

hist(Boston$ptratio, 25)
```
Yes, the median crime rate is 0.26, but the crime rate ranges from 0.01 to 88.98. Most cities have low crime rates, and the 5% of suburbs with the highest crime rate have crime rates larger than 15.78.

Tax rate is similar in that the median tax rate is 330, but there is a big jump between suburbs with lower tax rates and suburbs with tax rates higher than 650.

The pupil-teacher ratio distribution among the suburbs is pretty evenly distributed with the exception of a peak of suburbs with a pupil-teacher ratio of ~20.

e)
```{r}
sum(Boston$chas)
```
35 surburbs out of 506 bound the Charles River

f)
```{r}
median(Boston$ptratio)
```
The median pupil-teacher ratio is 19.05.

g)
```{r}
subset(Boston, medv == min(Boston$medv))
highlight <- function(x, value, col.value, col=NA, ...){
   hst <- hist(x, ...)
   idx <- findInterval(value, hst$breaks)
   cols <- rep(col, length(hst$counts))
   cols[idx] <- col.value
   hist(x, col=cols, ...)
}
highlight(Boston$crim, Boston$crim[Boston$medv == min(Boston$medv)], "red")
highlight(Boston$zn, Boston$zn[Boston$medv == min(Boston$medv)], "red")
highlight(Boston$indus, Boston$indus[Boston$medv == min(Boston$medv)], "red")
highlight(Boston$chas, Boston$chas[Boston$medv == min(Boston$medv)], "red")
highlight(Boston$nox, Boston$nox[Boston$medv == min(Boston$medv)], "red")
highlight(Boston$rm, Boston$rm[Boston$medv == min(Boston$medv)], "red")
highlight(Boston$age, Boston$age[Boston$medv == min(Boston$medv)], "red")
highlight(Boston$dis, Boston$dis[Boston$medv == min(Boston$medv)], "red")
highlight(Boston$rad, Boston$rad[Boston$medv == min(Boston$medv)], "red")
highlight(Boston$tax, Boston$tax[Boston$medv == min(Boston$medv)], "red")
highlight(Boston$ptratio, Boston$ptratio[Boston$medv == min(Boston$medv)], "red")
highlight(Boston$black, Boston$black[Boston$medv == min(Boston$medv)], "red")
highlight(Boston$lstat, Boston$lstat[Boston$medv == min(Boston$medv)], "red")
```
High crime, low proportion of large lots, large proportion of non-retail business, not bound by Charles River, larger nitrogen oxide concentration, oldest homes, closest to employment centers, most accessible to radial highways, higher property taxes, higher student-teacher ratios, higher proportion of blacks, and higher proportion of lower status. 

```{r}
nrow(subset(Boston, rm > 7))
nrow(subset(Boston, rm > 8))
summary(subset(Boston, rm > 8))
summary(Boston)
```
There are 64 suburbs with more than 7 rooms per dwelling, and 13 suburbs with more than 8 rooms per dwelling. 

The suburbs with more than 8 rooms per dwelling have less crime, higher median value, and lower lstat.

##Chapter 3, #15